---
title: "00_eda"
author: "Olga"
date: "10/17/2020"
output: html_document
---

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(kableExtra)
library(dplyr)
library(stringr)
```
# train.csv and submission.csv
```{r}
train <- fread("train/train.csv")
submission <- fread("sample_submission.csv")
```

Train size is `r dim(train)`

Balanced classes:

- Number of 1's in train `r sum(train$target)` (`r round(100*mean(train$target))`%).
- Number of 0's in train `r sum(!train$target)` (`r round(100*mean(!train$target))`%).



Number of unique phone_id's in train  `r uniqueN(train$phone_id)`

Number of unique phone_id's in submission  `r uniqueN(submission$Id)`

Number of unique phone_id's in submission  which are in train is `r sum(submission$Id %in% train$phone_id)`

For `r sum(!(submission$Id %in% train$phone_id))` clients there is no info in the train  dataset.


```{r}
records_per_phone <- train[, .N, by = phone_id]
ggplot(records_per_phone, aes(x=N)) + geom_histogram() + 
  ggtitle('Distribution of number of entries per phone ids')
```


```{r}
records_per_month <- train[, .N, by = order_completed_at ]
ggplot(records_per_month, aes(x=order_completed_at, y=N)) + 
  geom_bar(stat = "identity") +
  ggtitle("Number of records per month")
```
# shipments

Add phone_id to shipments
```{r}
shipments_1 <- fread("shipments/shipments2020-01-01.csv", encoding = 'UTF-8')
shipments_2 <- fread("shipments/shipments2020-03-01.csv", encoding = 'UTF-8')
shipments_3 <- fread("shipments/shipments2020-04-30.csv", encoding = 'UTF-8')
shipments_4 <- fread("shipments/shipments2020-06-29.csv", encoding = 'UTF-8')
addresses <- fread("misc/addresses.csv")

all_shipments <- rbind(shipments_1, shipments_2, shipments_3, shipments_4)
remove( list = c('shipments_1', 'shipments_2', 'shipments_3', 'shipments_4'))


sum(unique(all_shipments$user_id) %in% unique(train$phone_id))

user_id_ship_address_id = unique(all_shipments[, list(user_id, ship_address_id)])
#addresses$ship_address_id <- addresses$id
#addresses[, id := NULL]


uniqueN(train$phone_id)
uniqueN(addresses$phone_id)
sum(unique(train$phone_id) %in% unique(addresses$phone_id))
sum(unique(submission$Id) %in% unique(addresses$phone_id))

setkey(all_shipments, ship_address_id)
setkey(addresses, id)
addresses <- unique(addresses)
all_shipments_phone <- merge(all_shipments, addresses, by.x = 'ship_address_id',
                             by.y =  'id', all.x = TRUE)
```


Number of shipments per phone_id



```{r}
all_shipments_phone[, .N, by = phone_id] %>%
  ggplot(aes(x=N)) + geom_histogram(bins = 20) + 
  scale_x_log10() +
  scale_y_log10() +
  ggtitle('Distribution of number of shipments per phone id')
```

```{r}
table(all_shipments_phone[, .N, by = phone_id]$N)
```

## Retailers


```{r}
N <- 10
barplot(rev(sort(table(all_shipments$retailer)))[1:N], horiz = TRUE,las=1,
        main = paste0('Top ', N,  ' retailers'))
```

## Cities

```{r}
barplot(rev(sort(table(all_shipments$s.city_name)))[1:N], horiz = TRUE,las=1,
        main = paste0('Top ', N,  ' Cities'))

```

# Generate sex, age features

```{r}
all_phone_ids = unique(c(train$phone_id, submission$Id))

ids <- data.frame(phone_id = all_phone_ids)
setDT(ids)
ids <- merge(ids, addresses, by.x = 'phone_id', by.y = 'phone_id', all.x = TRUE)
ids$ship_address_id  <- ids$id
ids[, id := NULL]

ids <- merge(ids, all_shipments[, list(ship_address_id, user_id) ], by= 'ship_address_id', all.x = TRUE)
fwrite(ids, '01_preprocessed_data/user_id__phone_id__ship_address_id.csv')

user_profiles <- fread( 'misc/user_profiles.csv' )
ids[, ship_address_id := NULL]
ids <- unique(ids)
ids <- merge(ids, user_profiles, by='user_id', all.x = TRUE)
ids$age <-  as.numeric( as.Date("2020-01-01", "%Y-%m-%d") - as.Date(ids$bdate, "%Y-%m-%d") )/365
ids[, bdate := NULL]
fwrite(unique(ids), '01_preprocessed_data/user_id__phone_id__gender_age.csv')
```


# Prepare  train dataset

```{r}
ids_not_in_train <- submission$Id[!(submission$Id %in% train$phone_id)]
fwrite(data.frame(ids_not_in_train = ids_not_in_train),
       '01_preprocessed_data/submission_phone_ids_not_in_train.csv')

```

```{r}

train_v1 <- copy(train)
train_v1$month <- str_replace( train_v1$order_completed_at, "2020-", "")
train_v1$month <- as.numeric(train_v1$month )

train_v1_all <- train_v1
train_v1 <- train_v1[month < 7]

gender_age <- unique(ids[, list(phone_id, gender, age) ])
gender_age <- gender_age[!is.na(gender)]
gender_age <- gender_age[!is.na(age)]


temp <-  train_v1[, list(months_with_order = sum(target),
             months_without_order = sum(!target),
             last_order_month = max(month*target),
             first_order_month = min(month*target)), by = phone_id]

phone_id_label  <- train[order_completed_at  == '2020-07', list(phone_id, target)]

train_v1 <- merge(temp, phone_id_label, by = 'phone_id', all.x =  TRUE)


train_v1 <- merge(train_v1, gender_age, by = 'phone_id', all.x =  TRUE)



```


# Prepare submission 
```{r}
submission

train_v1_all <- train_v1_all[, list(months_with_order = sum(target),
             months_without_order = sum(!target),
             last_order_month = max(month*target),
             first_order_month = min(month*target)), by = phone_id]


train_v1_all <- merge(train_v1_all, gender_age, by = 'phone_id', all.x =  TRUE)
train_v1_all <- unique(train_v1_all)


submission_feat <- merge(submission, train_v1_all, by.x = 'Id',  by.y = 'phone_id', all.x = TRUE)



submission_feat <- submission_feat[!duplicated(Id)]
submission_feat$number <- seq(1, nrow(submission_feat))
  
  
submission_feat <- submission_feat[order(number)]
submission_feat[, Predicted := NULL]
submission_feat[, number := NULL]


```



```{r}
number_shipments <- all_shipments_phone[, list(n_orders = .N), by = phone_id ]

checks <- all_shipments_phone[, list(avg_check = mean(total_cost),
                                               max_check = max(total_cost),
                                               sum_spent = sum(total_cost),
                                     last_check = total_cost[.N],
                                     max_rate =  max(rate),
                                     was_rate_1 =  1*any(rate == 1),
                                     last_rate = rate[.N] ),
                                        by = phone_id ]



submission_feat <- merge(submission_feat, number_shipments, by.x = 'Id',  by.y = 'phone_id', all.x = TRUE)
train_v1 <- merge(train_v1, number_shipments, by = 'phone_id', all.x =  TRUE)

submission_feat <- merge(submission_feat, checks, by.x = 'Id',  by.y = 'phone_id', all.x = TRUE)
train_v1 <- merge(train_v1, checks, by = 'phone_id', all.x =  TRUE)


fwrite(train_v1, '01_preprocessed_data/train_v12.csv')
fwrite(submission_feat, '01_preprocessed_data/submission_features.csv')

```


